Usability of Modern Display Technologies

### 2014-11-20 mental-model-paper/paper2.tex
Inline comments and feedback for HW 1021.

\abstract{
This study aims to examine how interaction design concepts specifically map to the usability of modern display technology. This paper discusses the limitation of the lack of tactile feedback for touchscreen devices as well as how this limitation affects usability. This paper examines some current proposed solutions for making touch surfaces truly tangible. This limitation and each proposed solution has a variety of implications with respect to learnability, efficiency, satisfaction and errors.

}

@@

Electronic screens are devices through which information is transferred between the user and the interface. Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922.\cite{Cathode}} they are still diversifying in a multitude of ways to tackle the limitations that have presented themselves over the years.\cite{Eisenberg} Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks. For example, \textit{focus plus context screens}, which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents too large to fit standard screens.\cite{Baudisch} Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields.\cite{Hirsch}

@@

\section{Background/ Prior Work/ Literary Review}
\label{background}
\subsection{Limitations/Proposed Solutions}
There exists a fair amount of previous work pertaining to the usability of touch sensitive screens. The literature often evaluates usability by examining the limitations of these screen devices, as well as the proposed solutions for the limitations. One frequently discussed limitation is how typing on flat surfaces---with no physical keys to guide the fingers---requires heightened visual attention. A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed. \cite{Tinwala:2010:ETE:1868914.1868972} The lack of tactile features not only diverts attention from thoughts being expressed, but moreover, it makes the devices difficult to use for Individuals with Blindness or Severe Visual Impairment (IBSVI). A study from Virginia Polytechnic Institute and State University examined how touch screens---even those that include features such as  \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}}---do not provide sufficient support for IBSVI. These users were only able to ``develop a spatial mental model for the interface or the screen through dead reckoning'' since there was an ``absence of any landmark other than the boundary of the device.'' \cite{El-Glaly:2013:TTF:2460625.2460665}

@@

Another limitation of touchscreen devices that is examined in the literature is how the human finger as a pointing device has ``very low resolution.''\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger. A notable technique was created by Schneiderman, Sears, and colleagues to specifically address this problem.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen. The user drags the cursor to a desired target and then lifts the finger to select the targeted object. Although this technique generally allowed a user to perform precise movements with less errors, the user's efficiency tended to suffer.\cite{Sears}

@@

\subsection{Advantages}
Previous work on the usability of touchscreen display devices not only discusses limitations and proposed solutions, but also examines advantages of touchscreen displays. Ben Schneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses}, suggests a number of reasons why touchscreen devices can be advantageous: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc. \cite{73754} Many experiments have been conducted that provide evidence supporting Schneiderman's reasons. For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} Schneiderman's assertion that touchscreen display devices are genertally more learnable was also supported by an experiment that examined the learnability of a touchscreen electronic medical record system.\cite{Douglas:2011:SUL:2029976.2029990}

@@

Previous research has explored techniques that make touch screens more accessible to IBSVI by modifying the hardware. This paper previously mentioned one such technique when discussing a physical overlay. (See Section~\ref{background}) However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people.'' \cite{Kane:2011:AOI:2047196.2047232}  The study asserted that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.

@@

>>>>

\begin{abstract} % JD: My bad; this was the right way.
This study aims to examine how interaction design concepts specifically map to the usability of modern display technology. This paper discusses the limitation of the lack of tactile feedback for touchscreen devices as well as how this limitation affects usability. This paper examines some current proposed solutions for making touch surfaces truly tangible. This limitation and each proposed solution has a variety of implications with respect to learnability, efficiency, satisfaction and errors. 
\end{abstract}

@@

Electronic screens are devices through which information is transferred between the user and the interface. Although electronic displays have been around for roughly a century,\footnote{One of the earliest electronic displays, the cathode ray tube (CRT),  was first made commercial in 1922 \cite{Cathode}.} they are still diversifying in a multitude of ways to tackle the limitations that have presented themselves over the years \cite{Eisenberg}. Screen devices have been changing in size, shape, flexibility, and resolution in order to improve usability for specific tasks. For example, \textit{focus plus context screens}, which are wall-size low-resolution displays with an embedded high-resolution display region, are currently a proposed solution to working with visual documents too large to fit standard screens \cite{Baudisch}. Another kind of screen that is emerging is the \textit{BiDirectional (BiDi) Screen}, which is a thin, depth-sensing liquid-crystal display (LCD) that allows for 3D interaction using light fields \cite{Hirsch}.
% JD: I won't change it for the rest of the document, but just note that you want your cites *before*
%     the periods of sentences.

@@

\section{Background/ Prior Work/ Literary Review}
\label{background}
\subsection{Limitations/Proposed Solutions}
There exists a fair amount of previous work pertaining to the usability of touch sensitive screens
\cite{Tinwala:2010:ETE:1868914.1868972,El-Glaly:2013:TTF:2460625.2460665}. % JD: etc.
% JD: A sentence like the above typically concludes with a list of citations---representing the
%     aforementioned "fair amount of previous work." :)  Then these citations are cherry-picked
%     and discussed in further detail..the way you do next.
The literature often evaluates usability by examining the limitations of these screen devices, as well as the proposed solutions for the limitations. One frequently discussed limitation is how typing on flat surfaces---with no physical keys to guide the fingers---requires heightened visual attention. A study by Hussain Tinwala and Scott MacKenzie suggests that since the visual demand on the user is increased, concentration is diverted from the thoughts being expressed. \cite{Tinwala:2010:ETE:1868914.1868972} The lack of tactile features not only diverts attention from thoughts being expressed, but moreover, it makes the devices difficult to use for Individuals with Blindness or Severe Visual Impairment (IBSVI). A study from Virginia Polytechnic Institute and State University examined how touch screens---even those that include features such as  \textit{VoiceOver}\footnote{\textit{VoiceOver} for OS X tells users what is on the screen and walks users through actions.\cite{VoiceOver}}---do not provide sufficient support for IBSVI. These users were only able to ``develop a spatial mental model for the interface or the screen through dead reckoning'' since there was an ``absence of any landmark other than the boundary of the device.'' \cite{El-Glaly:2013:TTF:2460625.2460665}

@@

Another limitation of touchscreen devices that is examined in the literature is how the human finger as a pointing device has ``very low resolution.''\cite{Albinsson} With mobile touch screens in particular, research has illustrated the difficulty in pointing at targets that are often smaller than the width of the user's finger. A notable technique was created by Shneiderman, Sears, and colleagues to specifically address this problem.\cite{Sears} Their basic technique provides a cursor above the user's finger tip with a fixed offset when touching the screen. The user drags the cursor to a desired target and then lifts the finger to select the targeted object. Although this technique generally allowed a user to perform precise movements with less errors, the user's efficiency tended to suffer.\cite{Sears}

@@

\subsection{Advantages}
Previous work on the usability of touchscreen display devices not only discusses limitations and proposed solutions, but also examines advantages of touchscreen displays. Ben Shneiderman's essay, \textit{Touch Screens Now Offer Compelling Uses}, suggests a number of reasons why touchscreen devices can be advantageous: they are forms of direct manipulation, they are easy to learn, they are the fastest pointing device, no extra workspace is required to use them, etc. \cite{73754} Many experiments have been conducted that provide evidence supporting Shneiderman's reasons. For example, the results from an experiment by Robert Hardy and Enrico Rukzio suggest that finger interaction is, in fact, faster than alternative pointing devices in most situations.\cite{Hardy:2008:TIT:1409240.1409267} Shneiderman's assertion that touchscreen display devices are genertally more learnable was also supported by an experiment that examined the learnability of a touchscreen electronic medical record system.\cite{Douglas:2011:SUL:2029976.2029990}

@@

Previous research has explored techniques that make touch screens more accessible to IBSVI by modifying the hardware. This paper previously mentioned one such technique when discussing a physical overlay (See Section~\ref{background}).
% JD: Parentheticals like the one above also typically precede the period.
However, a recent study articulated how hardware modifications ``may be expensive to install, may limit the flexibility of the underlying software (by imposing physical structures), and may interfere with use by sighted people.'' \cite{Kane:2011:AOI:2047196.2047232}  The study asserted that for these reasons, techniques that rely upon hardware modifications have not been widely adopted.

% JD: You don't do this very often, but have done it enough times to get noticed:
%     sometimes you quote your references directly, but the quote itself does not
%     seem particularly compelling.  i.e., You would have done just as well to
%     paraphrase the author.  The reason this is preferred is that it gives your
%     writing a more cohesive voice---yours.  Unless a quote has such compelling
%     word choice and phrasing that its point loses value when paraphrased, things
%     generally read more smoothly if you don't disrupt the flow with quotes.
%
%     And note that you are still citing your source, so the paraphrasing does not
%     constitute any form of unfair use.

@@

% JD: Nicely done!  Please consider entering this in ULRA.  Additional figures might be
%     needed, I think, to communicate with a broader audience, but overall this is a
%     well-investigated, well-executed piece of work.
